---
title: "clean notebook"
author: "Kihel Abderraouf Saleh"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse,
               tidymodels,
               keras3,
               tensorflow,
               bonsai,
               ChainLadder,
               readxl,
               janitor,
               lubridate)
```

```{r}
data <- read_excel("data.xlsm")
```

```{r}
data <- data %>% clean_names()
```

```{r}
data %>% mutate(date_survenance = dmy(date_survenance)) -> data
```


```{r}
data %>% group_by(exercice) %>% arrange(exercice,date_survenance) %>% select(exercice,date_survenance,2:5,7) %>% mutate(year_of_sinistre = year(date_survenance)) ->boop

```

```{r}
boop %>% distinct() -> boop_clean
```

```{r}
boop_clean %>% 
  mutate(development_year = exercice - year_of_sinistre) %>% 
  group_by(year_of_sinistre, development_year) %>%
  summarise(total_reglement = sum(reglement), .groups = 'drop') %>% 
  arrange(development_year) %>% 
  pivot_wider(names_from = development_year, values_from = total_reglement, names_prefix = "Dev_Year_") %>% 
  arrange(year_of_sinistre) -> clean_data
```


# first branch Responsabilite civile 

```{r}
responsabilite_civile <- boop_clean %>% 
  filter(sous_branche == "Responsabilité Civile") %>% 
  mutate(development_year = exercice - year_of_sinistre) %>% 
  group_by(year_of_sinistre, development_year) %>%
  summarise(total_reglement = sum(reglement), .groups = 'drop') %>% 
  arrange(development_year) %>% 
  pivot_wider(names_from = development_year, values_from = total_reglement, names_prefix = "Dev_Year_") 

# the upper side will be filled with monte-carlo simulation in python.
```

# first model 

```{r}
responsabilite_civile_filled <- read_csv("respociviletrain.csv")
```

```{r}
responsabilite_civile_filled_split <- responsabilite_civile_filled %>% initial_split(prop = 0.75,strata = `Claim Amount`)
```

```{r}
responsabilite_civile_train <- responsabilite_civile_filled_split %>% training()
```

```{r}
responsabilite_civile_test <- responsabilite_civile_filled_split %>% testing()
```

```{r}
responsabilite_model <- rand_forest() %>% 
  set_engine("ranger") %>% 
  set_mode("regression")
```



```{r}
responsabilite_recipe <- responsabilite_civile_train %>% 
  recipe(`Claim Amount`~.) %>% 
  step_log(all_outcomes())
  
```

```{r}
responsabilite_wf <- workflow() %>% 
  add_model(responsabilite_model) %>% 
  add_recipe(responsabilite_recipe)
```

```{r}
responsabilite_fit <- responsabilite_wf %>% 
  last_fit(split = responsabilite_civile_filled_split)
```


```{r}
responsabilite_fit %>% collect_predictions() %>% mutate(.pred = exp(.pred),`Claim Amount` = exp(`Claim Amount`)) %>% 
  rmse(truth = `Claim Amount`,estimate = .pred) %>% mutate(mse = .estimate^2) -> responsabilite_metrics
```

prediction on the actual testing set : 


```{r}
responsabilite_real_test <- read_csv("respociviletest.csv") %>% select(2:4)
```

```{r}
responsabilite_fit %>% extract_fit_parsnip() %>% predict(responsabilite_real_test) %>% mutate(.pred = exp(.pred)) -> responsabilite_real_pred

```

predicting 2023 values :

```{r}
responsabilite_real_test %>% filter(Year == 2023) -> responsabilite_2023

responsabilite_fit %>% extract_fit_parsnip() %>% predict(responsabilite_2023) %>% mutate(.pred = exp(.pred)) -> responsabilite_pred_2023

responsabilite_pred_2023
```




## final matrix representation

```{r}
responsabilite_real_test2 <- responsabilite_real_test %>% mutate(`Claim Amount` = responsabilite_real_pred$.pred)

responsabilite_civile_filled %>% bind_rows(responsabilite_real_test2) %>% 
  pivot_wider(names_from = `Development Year`,values_from = `Claim Amount`) 
```


```{r}
responsabilite_civile_train 
responsabilite_civile_test2 <- responsabilite_real_test %>% mutate(`Claim Amount` = responsabilite_real_pred$.pred)

responsabilite_civile_train %>% bind_rows(responsabilite_civile_test2) -> responsability_table
```

# second branch Responsabilite civile 

```{r}
incendie <- boop_clean %>% 
  filter(sous_branche == "Incendie") %>% 
  mutate(development_year = exercice - year_of_sinistre) %>% 
  group_by(year_of_sinistre, development_year) %>%
  summarise(total_reglement = sum(reglement), .groups = 'drop') %>% 
  arrange(development_year) %>% 
  pivot_wider(names_from = development_year, values_from = total_reglement, names_prefix = "Dev_Year_") 

# the upper side will be filled with monte-carlo simulation in python.
```

# first model 

```{r}
incendie_filled <- read_csv("Incendietrain.csv")
```

```{r}
incendie_split <- incendie_filled %>% initial_split(prop = 0.75,strata = `Claim Amount`)
```

```{r}
incendie_train <- incendie_split %>% training()
```

```{r}
incendie_test <- incendie_split %>% testing()
```


```{r}
incendie_model <- rand_forest() %>% 
  set_engine("ranger") %>% 
  set_mode("regression")
```

```{r}
incendie_recipe <- incendie_train %>% 
  recipe(`Claim Amount`~.) %>% 
  step_log(all_outcomes())
```

```{r}
incendie_wf <- workflow() %>% 
  add_model(incendie_model) %>% 
  add_recipe(incendie_recipe)
```

```{r}
incendie_fit <- incendie_wf %>% 
  last_fit(split = incendie_split)
```

```{r}
incendie_fit %>% collect_predictions() %>% mutate(.pred = exp(.pred),`Claim Amount` = exp(`Claim Amount`)) %>% 
  rmse(truth = `Claim Amount`,estimate = .pred) %>% mutate(mse = .estimate^2) -> incendie_metrics
```

prediction on the actual testing data

```{r}
incendie_real_test <- read_csv("Incendietest.csv") %>% select(2:4)
```

```{r}
incendie_fit %>% extract_fit_parsnip() %>% predict(incendie_real_test) %>% mutate(.pred = exp(.pred)) -> incendie_real_preds
```

```{r}
incendie_real_test2 <- incendie_real_test %>% mutate(`Claim Amount` = incendie_real_preds$.pred)
```

predicting on 2023 data 


```{r}
incendie_real_test %>% filter(Year == 2023) -> incendie_2023
```


```{r}
incendie_fit %>% extract_fit_parsnip() %>% predict(incendie_2023) %>% mutate(.pred = exp(.pred)) -> incendie_pred_2023
```

```{r}
incendie_pred_2023
```


# original matrix format

```{r}
incendie_filled %>% bind_rows(incendie_real_test2) %>% pivot_wider(names_from = `Development Year`,values_from = `Claim Amount`)
```

```{r}
incendie_train
incendie_real_test$`Claim Amount` <- incendie_real_preds$.pred

incendie_train %>% bind_rows(incendie_real_test) -> incendie_table
```


# branch 3 , risque simple 

```{r}
risk <- boop_clean %>% 
  filter(sous_branche == "Risque simple") %>% 
  mutate(development_year = exercice - year_of_sinistre) %>% 
  group_by(year_of_sinistre, development_year) %>%
  summarise(total_reglement = sum(reglement), .groups = 'drop') %>% 
  arrange(development_year) %>% 
  pivot_wider(names_from = development_year, values_from = total_reglement, names_prefix = "Dev_Year_") 

# the upper side will be filled with monte-carlo simulation in python.
```

```{r}
risk_filled <- read_csv("risquesimpletrain.csv")
```
```{r}
risk_filled
```


```{r}
risk_split <- risk_filled %>% initial_split(prop = 0.75,strata = `Claim Amount`)
```

```{r}
risk_train <- risk_split %>% training()
```

```{r}
risk_test <- risk_split %>% testing()
```

```{r}
risk_model <- boost_tree() %>% 
  set_engine("lightgbm",objective = "tweedie",tweedie_variance_power = 1.5) %>% 
  set_mode("regression")
```


```{r}
risk_recipe <- risk_train %>% 
  recipe(`Claim Amount`~.) %>% 
  step_log(all_outcomes())
```


```{r}
risk_wf <- workflow() %>% 
  add_model(risk_model) %>% 
  add_recipe(risk_recipe)
```

```{r}
risk_fit <- risk_wf %>% last_fit(risk_split)
```

```{r}
risk_fit %>% collect_predictions() %>% mutate(.pred = exp(.pred),`Claim Amount` = exp(`Claim Amount`)) %>% rmse(estimate = .pred,truth = `Claim Amount`) %>% mutate(mse = .estimate^2) -> risk_metrics
```

prediction on the actual testing data 

```{r}
risk_real_test <- read_csv("risquesimpletest.csv") %>% select(2:4)
```

```{r}
risk_real_pred <- risk_fit %>% extract_fit_parsnip() %>% predict(risk_real_test)
```

```{r}
risk_real_pred %>% mutate(.pred = exp(.pred)) -> risk_real_pred
```

```{r}
risk_real_test2 <- risk_real_test %>% mutate(`Claim Amount` = risk_real_pred$.pred)
```

```{r}
risk_real_test %>% filter(Year == 2023) -> risk_2023
```

```{r}
risk_fit %>% extract_fit_parsnip() %>% predict(risk_2023) %>% mutate(.pred = exp(.pred)) -> risk_pred_2023
```

```{r}
risk_pred_2023
```






# original matrix 

```{r}
risk_filled %>% bind_rows(risk_real_test2) %>% pivot_wider(names_from = `Development Year`,values_from = `Claim Amount`)
```


```{r}
risk_train

risk_real_test$`Claim Amount` <- risk_real_pred$.pred

risk_train %>% bind_rows(risk_real_test) -> risk_table
```




# branch 4 cat-nat

```{r}
cat <- boop_clean %>% 
  filter(sous_branche == "CAT-NAT") %>% 
  mutate(development_year = exercice - year_of_sinistre) %>% 
  group_by(year_of_sinistre, development_year) %>%
  summarise(total_reglement = sum(reglement), .groups = 'drop') %>% 
  arrange(development_year) %>% 
  pivot_wider(names_from = development_year, values_from = total_reglement, names_prefix = "Dev_Year_") 

# the upper side will be filled with monte-carlo simulation in python.
```

```{r}
cat_filled <- read_csv("catnattrain.csv")
```

```{r}
cat_split <- cat_filled %>% initial_split(prop = 0.75)
```

```{r}
cat_train <- cat_split %>% training()
```

```{r}
cat_test <- cat_split %>% testing()
```



```{r}
cat_recipe <- cat_train %>% 
  recipe(`Claim Amount`~.) %>% 
  step_log(all_outcomes())
```


```{r}
cat_model <- linear_reg(penalty = 1e-4,mixture = 0.5) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet")
```

```{r}
cat_wf <- workflow() %>% 
  add_model(cat_model) %>% 
  add_recipe(cat_recipe)
```

```{r}
cat_fit <- cat_wf %>% last_fit(split = cat_split)
```

```{r}
cat_fit %>% collect_predictions() %>% mutate(.pred = .pred^2, `Claim Amount` = (`Claim Amount`)^2) %>% rmse(truth = `Claim Amount`,estimate = .pred) %>% mutate(mse = .estimate^2) -> cat_metrics
```

testing on real data

```{r}
cat_real_test <- read_csv("catnattest.csv") %>% select(2:4)
```

```{r}
cat_real_pred <- cat_fit %>% extract_fit_parsnip() %>% predict(cat_real_test) %>% mutate(.pred = exp(.pred))
```

```{r}
cat_real_pred2 <- cat_real_test %>% mutate(`Claim Amount` = cat_real_pred$.pred)
```


# original matrix format

```{r}
cat_filled %>% bind_rows(cat_real_pred2) %>% pivot_wider(values_from = `Claim Amount`,names_from = `Development Year`) 
```


```{r}
cat_train 
cat_real_test$`Claim Amount` <- cat_real_pred$.pred

cat_train %>% bind_rows(cat_real_test) -> cat_table
```

```{r}
responsability_table %>% bind_rows(incendie_table,risk_table,cat_table) 
```



```{r}
responsabilite_metrics %>% bind_rows(incendie_metrics,risk_metrics,cat_metrics) %>% pull(mse) -> models_mse
 
model_metrics <- tibble(category = c("Responsabilité civile","Incendie","Risque simple","CAT-NAT"),Model = c("Random Forest","Random Forest","Random Forest","Linear regression"),mse =models_mse)

model_metrics
```



```{r}

responsabilite_fit %>% extract_fit_parsnip() %>% butcher::butcher() %>% saveRDS("responsabilite_model.rds")
incendie_fit %>% extract_fit_parsnip() %>% butcher::butcher() %>% saveRDS("incendie_model.rds")
risk_fit %>% extract_fit_parsnip() %>% butcher::butcher() %>% saveRDS("risk_model.rds")
cat_fit %>% extract_fit_parsnip() %>% butcher::butcher() %>% saveRDS("cat_model.rds")
```

















































































































































































































